---
title: "saturate: A Complete User Guide"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{saturate: A Complete User Guide}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse  = TRUE,
  comment   = "#>",
  eval      = FALSE
)
```

*Qualitative Data Analysis for R* -- Version 0.1.0

# Part 1 -- Getting Started

## Chapter 1: What Is saturate?

saturate is an R package for qualitative data analysis. It gives you a
browser-based graphical interface -- a Shiny application -- for reading
interview transcripts, attaching codes to passages of text, organising those
codes into hierarchies and themes, comparing the work of multiple coders, and
exporting results for publication. You can also drive every operation from R
scripts if you prefer to work programmatically.

The name reflects a core concept in qualitative research: theoretical
saturation, the point at which new data stops producing new insights. saturate
helps you track that point systematically.

### Why saturate exists

For many years, the most widely used free qualitative analysis tool in R was
RQDA. It was archived by CRAN because it depended on an outdated version of the
GTK graphics library that could no longer be distributed reliably. This left a
gap in the R ecosystem for qualitative researchers.

saturate fills that gap. It stores all project data in a single DuckDB file,
which is portable, fast, and requires no database server. The Shiny interface
runs in your browser, so it looks and behaves consistently on Windows, macOS,
and Linux. And because every operation is also available as an R function, you
can automate imports, run reliability statistics, and integrate qualitative data
with quantitative analyses in a single script.

### The running example

Throughout this guide you will follow Researcher A, a qualitative researcher
studying remote working and wellbeing. Researcher A has conducted twelve
semi-structured interviews with knowledge workers across three organisations --
Organisation A, Organisation B, and Organisation C -- covering topics such as
work-life boundaries, isolation, productivity, management practices, and mental
health. Their transcripts are saved as plain-text files. They have a research
assistant, Researcher B, who will independently code a subset of the data so
they can calculate inter-rater reliability.

---

## Chapter 2: Installation

### Prerequisites

You need R version 4.1 or later. You do not need to install DuckDB, a database
server, or any other software separately -- all dependencies are handled by R's
package manager.

### Installing saturate

Install saturate from CRAN:

```{r}
install.packages("saturate")
```

Or install the development version from GitHub:

```{r}
# install.packages("pak")
pak::pak("thomaswood/saturate")
```

After installation, load the package in your R session:

```{r}
library(saturate)
```

### A note on updates

When a new version of saturate is released, reinstall from CRAN or GitHub. Your
project files (`.satdb`) are forward-compatible -- opening an old project in a
newer version of saturate automatically adds any new database tables without
altering existing data.

---

## Chapter 3: Creating and Managing Projects

### What a saturate project is

Every saturate study lives in a single file with the `.satdb` extension (older
projects use `.duckdb`, which is equally supported). This file is a DuckDB
database -- a self-contained, portable data store that holds your documents,
codes, codings, cases, journal entries, and every other piece of project data.
You can copy it, email it, back it up to cloud storage, or hand it to a
colleague, and the entire project travels with it.

### Creating a new project

In R, create a new project with `qc_new()`:

```{r}
proj <- qc_new(
  path  = "~/research/remote_work/remote_wellbeing.satdb",
  name  = "Remote Working and Wellbeing",
  owner = "Researcher A"
)
```

The `path` argument is the full file path where you want the project file to be
saved. The `name` and `owner` arguments are stored as project metadata and
displayed in the interface. Both default to sensible values (`name` defaults to
the filename, `owner` to the system user) if omitted.

After running `qc_new()`, the variable `proj` holds an open connection to the
project. You can pass it directly to `shiny_saturate()` to launch the interface,
or use it with the API functions described in Part 6.

### Where to save your project file

Choose a location on your local hard drive or a folder that syncs to cloud
storage (such as OneDrive, Dropbox, or Google Drive). Avoid saving the file
directly to a network drive that other users can access simultaneously -- DuckDB
uses file-level locking, and two sessions writing to the same file at the same
time will cause one of them to error.

### Opening an existing project

```{r}
proj <- qc_open("~/research/remote_work/remote_wellbeing.satdb")
```

`qc_open()` checks that the file exists and is a valid saturate project. If
another R session already has the file open, `qc_open()` will return an error
-- only one session can write to a project at a time.

### Closing a project

```{r}
qc_close(proj)
```

This flushes any pending writes and releases the file lock. Always call this
when you finish a session.

### Launching the graphical interface

```{r}
shiny_saturate(proj)
```

Your default browser will open automatically. The R console will show
`Listening on http://127.0.0.1:XXXX` -- this is a local web server running on
your own computer. Your data never leaves your machine.

To stop the app, press **Escape** or **Ctrl+C** in the R console.

### Launching without a project (the project launcher)

You can also call `shiny_saturate()` without a project argument:

```{r}
shiny_saturate()
```

This opens a **project launcher** screen in the browser. From there you can:

- **Open** an existing `.satdb` or `.duckdb` file by typing or browsing to its
  path.
- **Create** a new project by choosing a folder and entering a project name.

The launcher is convenient when you want to let a non-technical colleague open
or create a project without writing R code.

---

## Chapter 4: The saturate Interface

When you launch `shiny_saturate(proj)`, the browser window opens to the
**Documents** tab. Across the top of the screen is the navigation bar (navbar)
which gives you access to every section of the application.

### The navbar

| Tab | Purpose |
|---|---|
| **Documents** | Import, view, and manage your source documents |
| **Coding** | Read documents and attach codes to passages |
| **Compare** | Compare two coders' work side by side |
| **Codebook** | Create and manage codes, categories, and codebook snapshots |
| **Themes** | Organise codes into higher-level themes |
| **Query** | Search documents, explore co-occurrence, triangulate, visualise, word clouds |
| **Cases** | Manage participants or sites; set attributes; link documents |
| **Journal** | Write reflexive and analytical memos |
| **Members** | Member checking: generate summaries for participants to review |
| **Export** | Download data in various formats |
| **Audit** | View the full code change history and coding audit log |
| **Help** | Built-in reference documentation |

### The session area (top right)

**Coding As** text box: The name that will be attached to every coding you
create. It is populated automatically when you choose a profile (see Chapter 5).

**Blind** toggle: When on, you will not see any highlights from other coders
while you work in the **Coding** tab. This allows two or more coders to work
independently without being influenced by each other's choices.

**Settings** button: Opens the Settings panel.

**Project** button: Opens the Project panel, which has three tabs -- **Info**
(project metadata), **Split** (create a contributor copy for independent
coding), and **Merge** (import a contributor's work back into the master). See
Chapter 11.

### The profile gate

The first time you open a project in a new browser session, saturate displays a
splash screen titled **Choose a profile**. You must select or create a profile
before the main interface becomes available.

---

## Chapter 5: Settings and Profiles

### What a profile is

A profile is a named coder identity. When you select a profile, saturate sets
the **Coding As** text to that profile's name and applies that profile's saved
interface preferences. Every coding you apply and every memo you write is stored
with this coder name.

For Researcher A's study, they create a profile called "Researcher A" and their
research assistant creates one called "Researcher B".

### Managing profiles

Click the **Settings** button in the navbar. In the **Profile** section:

- **Creating a profile**: Type a name in "Create profile" and click
  **Create and switch**.
- **Switching profiles**: Choose from "Active profile" and click **Switch**.
- **Deleting a profile**: Choose a profile and click **Delete**. This does not
  delete any codings or memos created under that name.

### Display settings

| Setting | Description |
|---|---|
| **Colour profile** | Light, Dark, High contrast, High contrast dark, Ocean, or Warm |
| **Interface font** | System UI, Serif, Readable sans, or Monospace |
| **Interface size** | Scale from 90% to 125% |
| **Document font** | Font used in the document pane |
| **Document text size** | Scale from 85% to 150% |
| **Document line spacing** | Line height from 1.4 to 2.4 |
| **Document pane height** | Height in viewport units (48vh to 86vh) |
| **Table density** | Compact, Comfortable, or Roomy |
| **Highlight opacity** | How strongly code colours overlay the text |
| **Show line numbers** | Display line numbers in the document margin |
| **Reduce animation** | Disable transitions and animated effects |

### Text-to-speech settings

| Setting | Description |
|---|---|
| **TTS voice** | Browser voice to use for read-aloud (leave blank for system default) |
| **Reading speed** | Playback rate from 0.5x to 2.0x |

---

# Part 2 -- Building Your Dataset

## Chapter 6: The Documents Tab

### What a document is in saturate

In saturate, a **document** (also called a **source**) is any piece of text you
want to analyse -- interview transcripts, focus group transcripts, open-ended
survey responses, field notes, policy documents, or any other textual material.

Researcher A imports twelve documents -- one per interview -- each named with
the participant's pseudonym and organisation.

### Importing documents

Navigate to the **Documents** tab and click **Import**. Select one or more
files. Supported file types:

- Plain text (`.txt`)
- Microsoft Word documents (`.docx`)
- PDF files (`.pdf`) -- text-layer PDFs only; scanned PDFs require OCR first
- CSV files (`.csv`) -- the first column is treated as the text content

You can also click **Paste text** to type or paste document content directly
into the interface without needing a file.

### The documents table

| Column | Meaning |
|---|---|
| **Name** | The document title |
| **Type** | The document type (interview, focus group, etc.) |
| **Length** | Total character count |
| **Codings** | How many coded segments exist across all coders |
| **Coders** | How many distinct coder names have applied at least one coding |
| **Memo** | Whether a document memo has been written |

The **Coders** column tells you whether multiple team members have worked on the
same document -- relevant when you need to calculate inter-rater reliability.

### Document memos

Click a document row to open the viewer, then use the **Memo** controls to
record methodological notes about the source (e.g., "This participant spoke
quickly; transcript may have minor errors in turns 45-52").

### Deleting a document

Click the document's row and then click **Delete**. You will be asked to
confirm. The record is soft-deleted -- marked inactive rather than physically
erased. All codings attached to that document are also soft-deleted.

---

## Chapter 7: The Cases Tab

### What a case is

A **case** is the unit of analysis in your study -- typically a participant, but
it could be a site, an organisation, or any other bounded entity you are
comparing. Cases allow you to group documents together and attach structured
descriptive attributes that you can later use for triangulation.

In Researcher A's study, each case is one participant. They record attributes
including organisation, role level, age group, working arrangement, and months
remote.

### Creating a case

Navigate to the **Cases** tab. In the sidebar, type a name in the **Name**
field (e.g., "P01"), optionally add a **Memo**, and click **Add case**.

### Attributes sub-panel

Click a case row to open its detail panel. In the **Attributes** sub-panel,
type a variable name (e.g., `working_arrangement`) and a value
(e.g., `fully remote`), then click **Set attribute**.

Researcher A adds:

```
organisation        = Organisation A
role_level          = senior
age_group           = 40s+
working_arrangement = fully remote
months_remote       = 24
```

Use consistent variable names across all cases. Inconsistent naming (e.g.,
`organisation` for P01 but `org` for P02) breaks the triangulation analysis.

To delete an attribute, click the attribute row in the table and confirm.

### Documents sub-panel

In the **Documents** sub-panel, choose a document from the dropdown and click
**Link**. Click a linked document row to unlink it. A document can be linked to
multiple cases.

### Exporting case attributes

Click **Attributes CSV** at the top of the Cases tab. saturate generates a
wide-format CSV where each row is a case and each column is an attribute.

---

# Part 3 -- Coding

## Chapter 8: The Codebook Tab

### What codes are

A **code** is a short label you apply to a passage of text to indicate that the
passage represents a particular concept, theme, or phenomenon.

In Researcher A's study, early codes include `boundary_blurring`,
`isolation_felt`, `flexibility_valued`, `manager_unavailability`,
`home_interruptions`, and `digital_fatigue`.

### Creating a code

Click the **New Code** button. The creation form has these fields:

- **Name** (required): A short, distinctive label. Underscores recommended over
  spaces.
- **Colour**: Click the swatch to open a colour picker.
- **Definition**: A one-to-two sentence explanation of what this code means.
  Critical for team consistency.
- **Criteria**: Notes on when to apply this code versus a similar one.
- **Memo**: Free-text notes about the code's development or status.
- **Parent code**: Select a parent to place this code within a hierarchy.
- **Level**: Analytic level (e.g., `descriptive`, `interpretive`).
- **Orientation**: Theoretical orientation (e.g., `phenomenological`).
- **Weight**: Optional numeric weight for the code.

### Code hierarchy

saturate supports parent-child hierarchies. Researcher A creates a parent code
`wellbeing_impact` and beneath it: `isolation_felt`, `digital_fatigue`,
`anxiety_work_performance`, and `positive_affect_flexibility`. Child codes
appear indented beneath their parent in the codebook table.

A code can only have one parent, but a parent can have any number of children.

### Categories

A **category** is a lateral grouping of codes that cuts across the hierarchy.
Use categories to group codes from several branches under a shared theme (e.g.,
"Management Relationships" could group codes from different branches that all
relate to manager behaviour).

### Importing and exporting a codebook

Click **Import** to load a codebook from CSV. Expected columns: `name`,
`definition`, `criteria`, `memo`, `parent`, `colour`, `weight`. Any missing
columns are treated as empty.

Click **Export** to download a CSV of all codes and their metadata. Include this
as a supplementary file in publications.

### The Validate button

Click **Validate** to run a consistency check. saturate checks for:

- **Orphan parents**: A code lists a parent that does not exist.
- **Circular hierarchy**: Code A is an ancestor of Code B while B is an
  ancestor of A.
- **Missing definitions**: Codes with no definition text.
- **Unused codes**: Codes created but not yet applied to any passage.

### The Snapshots button

Click **Snapshots** to take a named, timestamped record of your codebook at a
given point in time. Give it a meaningful name (e.g., "Post Phase 1 fieldwork").

To compare two snapshots, select both from the list and click **Diff**. saturate
shows which codes were added, removed, or modified between the two points. This
documents how your interpretive framework evolved -- important for rigorous
reporting.

### Code lifecycle: deprecating and merging codes

As your codebook matures you will sometimes need to retire a code or fold it
into another. Two operations support this:

- **Deprecate**: Marks a code as no longer in active use. It remains visible in
  history and its codings are preserved, but it is excluded from the code picker
  during coding.
- **Merge into**: Reassigns all codings from one or more source codes to a
  target code, then soft-deletes the sources. Use the **Merge into...** button
  in the codebook form, or `qc_merge_codes()` in the API.

---

## Chapter 9: The Coding Tab

### Opening a document for coding

Navigate to the **Coding** tab. Select a document from the document selector.
The transcript text appears in the main coding pane with any existing
highlights.

### Selecting text to code

Click and drag to select the text you want to code. Release the mouse button. A
**code picker** dialogue appears -- a searchable list of all codes in your
codebook. Type to filter by name, then click a code to apply it. The selection
is immediately highlighted in that code's colour.

To apply multiple codes to the same passage, select the same text again and
choose a different code.

### How overlapping codes work

Overlapping codes are handled correctly in saturate. The highlighting engine
calculates colour bands in R before sending the HTML to the browser, rather than
relying on the browser to manipulate the document structure. This means you can
have Code A covering an entire sentence, Code B covering the first half, and
Code C covering the last three words, and all three highlights render correctly.

### Blind mode

The **Blind** button in the navbar hides other coders' highlights so you see
only your own work. This allows independent coding without being anchored by a
colleague's existing highlights.

Researcher A asks Researcher B to turn Blind mode on before coding any document,
ensuring their codes reflect their independent interpretation. After both have
coded a document, Researcher A turns Blind mode off to compare their work in the
**Compare** tab.

### Removing a coding

Click a highlighted passage to see a context menu with the code name and a
**Remove** button. Removing a coding does not affect other codes on the same
passage.

### Confidence

When applying a code, you can optionally set a **confidence** value (0-100).
The default is 100. Use lower values for genuinely ambiguous passages.
Confidence values can be retrieved via the API for sensitivity analysis.

### Text-to-speech (read aloud)

The coding pane includes a **read-aloud** control bar beneath the document
selector. Click **play** (the triangle button) to begin reading from the current
scroll position. The document scrolls to follow along and the text cursor tracks
the reading position.

You can also start reading from a specific point by clicking anywhere in the
document text, or by selecting a passage -- playback begins from the start of
your selection.

Click **pause** to suspend reading; click **play** again to resume. Click the
square **stop** button to end the session and return to the beginning.

Voice and speed are controlled in the **Settings** panel under
**Text-to-speech settings**. The voices available depend on your operating
system and browser.

---

## Chapter 10: Auto-Coding (R API)

Auto-coding is available through the R API only. It applies a code
automatically to every passage that matches a regular expression.

```{r}
# Apply "mentions_flexibility" to every sentence containing "flexible"
qc_auto_code(
  proj,
  code_id     = 12,
  pattern     = "flexib",
  ignore_case = TRUE
)
```

Review auto-codings after applying them -- they are a scanning tool, not a
substitute for analytical reading. You can remove individual auto-codings in the
GUI.

---

## Chapter 11: Multi-Coder Workflows: Split and Merge

### Overview

saturate supports file-based collaboration: one researcher holds the **master**
project file, creates **contributor copies** for team members, and consolidates
everyone's work back into the master when analysis is complete. No shared server
or network drive is required -- files are exchanged by email, cloud storage, or
any other file-transfer method.

The workflow has three steps:

1. **Split**: The lead researcher creates one contributor copy per team member
   -- each a complete, self-contained `.satdb` file -- and distributes them.
2. **Code**: Each team member opens their copy and codes independently, then
   returns the file.
3. **Merge**: The lead researcher uploads each returned file. saturate adds any
   new codes, codings, themes, and memos to the master automatically, skipping
   items that are already there.

### Splitting a project

Click the **Project** button in the navbar and select the **Split** tab.

**Documents to include:**

- **All documents** -- every active document is copied to the contributor file.
- **Select documents** -- choose specific documents from the list. Use this when
  you want Researcher B to code only a subset of documents for reliability
  checking.

**Include existing codings:** If unchecked (the default), the contributor file
contains only the codebook and documents -- the coder starts fresh. Check this
box only when the coder needs to continue or review existing work.

Click **Download split project (.satdb)**. Your browser saves a `.satdb` file
named after the project and today's date. Send this file to the coder.

Researcher A creates a contributor file for Researcher B containing Documents 3,
7, and 11 with **Include existing codings** unchecked -- Researcher B will code
those three documents independently, producing a separate set of codings for
inter-rater reliability analysis.

### What is in a contributor file

The contributor file contains:

- The complete codebook (all active codes, categories, hierarchies, code
  relations)
- All cases and case attributes
- All themes and theme-code links
- The selected documents and their metadata
- Existing codings only if **Include existing codings** was checked
- Project name, owner, and settings

What is **not** copied: member checks, codebook snapshots, and the project lock
state.

### Merging a contributor file

When the coder returns their file, click the **Project** button and select the
**Merge** tab.

Click **Choose contributor file** and upload the returned file. Then click
**Preview**.

saturate scans the contributor file and shows what would be imported before
anything is changed:

```
Would import:
 * 2 new codes
 * 0 new documents
 * 47 codings  (3 duplicates will be skipped)
 * 0 new themes
 * 1 new memo
```

Under **On conflict**, choose what happens if the same coding already exists in
the master:

- **Skip existing** (default): leave the master's version untouched.
- **Replace existing**: soft-delete the master's version and insert the
  contributor's.

Click **Merge into project** to confirm. The master project is updated
immediately.

### What gets merged

| Data type | Merged? | How duplicates are detected |
|---|---|---|
| Codes | Yes -- new codes only | Matched by name (case-insensitive) |
| Code categories | Yes -- new only | Matched by name |
| Documents | Yes -- new only | Matched by content hash, then by name |
| Codings | Yes | Same document + code + start + end + coder |
| Themes | Yes -- new themes only | Matched by name |
| Project memos | Yes -- new only | Same content + author + type |
| Member checks | **No** | Always kept in the master only |
| Codebook snapshots | **No** | Not copied or merged |

### Merging multiple contributor files

Run the Merge step once per contributor file. Subsequent merges deduplicate
correctly -- if two contributors applied the same coding independently, it is
added on the first merge and skipped on the second.

### Using the R API

```{r}
# Split: create a contributor copy with Documents 3, 7, and 11
contrib <- qc_split_project(
  proj,
  path       = "~/collab/contrib_B_2026-04.satdb",
  source_ids = c(3L, 7L, 11L)
)
qc_close(contrib)

# Merge: import the returned file into master
result <- qc_merge_project(
  proj,
  contributor_path = "~/collab/contrib_B_returned.satdb"
)
result$codings_added  # new codings imported
result$codings_skip   # duplicate codings skipped
result$codes_added    # new codes propagated from contributor
```

To import codings from a specific coder only:

```{r}
qc_merge_project(
  proj,
  contributor_path = "~/collab/contrib_B_returned.satdb",
  coders           = "Researcher B"
)
```

---

# Part 4 -- Analysis

## Chapter 12: The Compare Tab

### What the Compare tab is for

The **Compare** tab places two coders' work on the same document side by side,
highlights where they agree and disagree, and calculates inter-rater reliability
statistics.

### Selecting document and coders

Navigate to the **Compare** tab. Choose the **Document**, **Coder A**, and
**Coder B** from the selectors at the top.

### The differences card

The **Differences** card lists every passage where the two coders' work
diverges:

- Coder A applied a code that Coder B did not.
- Coder B applied a code that Coder A did not.
- Both applied a code to overlapping (but not identical) character ranges.

Click a row to scroll the document view to that passage. Use these differences
as the basis for discussion and codebook refinement, and document your decisions
in the Journal.

### The reliability card

When in **coders** mode, a **Reliability** card appears showing:

- **Cohen's kappa per code**: How much better than chance the two coders agreed
  on each code.
- **Mean kappa**: Average across all codes.
- **Percentage agreement**: How often both coders made the same decision.
- **N pairs**: Number of passage-code pairs compared.

Kappa values are colour-coded: green (> 0.61), yellow (0.41-0.60),
red (< 0.40). Codes with poor kappa usually indicate an ambiguous definition.

---

## Chapter 13: The Themes Tab

### What a theme is

A **theme** is a higher-level interpretive structure synthesising several codes
and categories. Where a code is an observation ("this passage is about boundary
blurring"), a theme is an argument ("remote workers experience boundary blurring
primarily through temporal rather than spatial mechanisms").

### Creating a theme

Navigate to the **Themes** tab. Click **New Theme**. Give the theme a **Name**
and write a substantive **Description**. Researcher A creates "Temporal Boundary
Erosion": *"Remote workers describe the boundaries of their workday becoming
progressively undefined, with evening and weekend work normalised through
always-on digital communication."*

### Linking codes and categories to a theme

Click a theme in the table to open its detail panel. Under **Linked codes**, add
codes from your codebook. You can also link entire categories.

### Theme excerpts

The **Excerpts** sub-panel shows every coded passage from every document tagged
with any of the theme's linked codes -- a consolidated view across all
participants and documents. Filter by coder, document, or case attribute to
explore how the theme manifests in different subgroups.

### Exporting a themes report

Click **Export themes report**. saturate generates a structured document with
one section per theme, containing the description, linked codes, and all
associated excerpts with source documents and coder names.

---

## Chapter 14: The Query Tab

The **Query** tab consolidates eight analytical panels in one place.

### Coded segments panel

Retrieve all codings matching a combination of codes, documents, and coders.
Download the results as CSV. This is the fastest way to compile all passages
tagged with a particular code across your entire corpus.

### Full-text search panel

Type a search term in the **Search** box. saturate searches the full text of
all documents and returns every matching passage with surrounding context.

**Options:**

- **Regex**: Use regular expression patterns (e.g., `\bflexib\w*\b` matches
  "flexible", "flexibility", "flexibly").
- **Case-insensitive**: Usually best left on.
- **Context window**: Characters of surrounding text shown with each match.

Click **CSV** to download all results.

### Co-occurrence panel

Shows how frequently pairs of codes appear together in the same document or the
same overlapping passage. Useful for identifying codes that consistently travel
together and may belong in the same theme.

### Saturation curve panel

Plots the cumulative number of distinct codes introduced as documents are
processed in order. A curve that flattens provides empirical evidence of
theoretical saturation. See also `qc_saturation_curve()` in the R API.

### Triangulation panel

Cross-tabulate your codes against case attributes to reveal patterns across
participant groups. Select codes or categories and choose a **Case attribute**
to cross-tabulate against (e.g., `working_arrangement`).

Researcher A finds that `boundary_blurring` appears far more frequently in
fully-remote participants than hybrid participants, while `flexibility_valued`
appears at similar rates across both groups.

Click **CSV** to export the triangulation table.

### Cross-tabulation panel

Produce a code-by-document frequency table. Each cell contains the number of
codings of a given code in a given document. Useful for reporting patterns
across cases.

### Graph panel

The **Graph** panel renders a network visualisation with three modes selectable
from the dropdown:

- **Co-occurrence network**: Codes as nodes; edges connect codes that appear
  together in the same passage.
- **Bipartite network**: Documents and codes as nodes; edges represent coding
  links.
- **Similarity network**: Documents as nodes; edges connect documents that share
  codes.

Drag nodes to rearrange, zoom in/out, and hover to see counts.

### Word cloud panel

Generate a word cloud from your corpus or from the excerpts of a specific code.
Two source options are available:

- **All codes -- sized by coding count**: Each code label is shown, sized by
  how many times it has been applied.
- **Single code -- words in excerpts**: Shows the most frequent words found
  within all passages tagged with a chosen code.

Click **Generate** to render the cloud. Click **Download PNG** to save it.

---

## Chapter 15: Saturation Analysis (R API)

Theoretical saturation is the point at which new data stops producing new
codes. Demonstrating saturation is increasingly expected in published qualitative
research.

```{r}
curve <- qc_saturation_curve(proj)
plot(curve)
```

The function works through your documents in import order and counts how many
new codes appear with each successive document. The curve typically rises
steeply for the first few documents and then flattens.

Researcher A's curve flattens after document 9, providing empirical evidence of
saturation included as a supplementary figure. The saturation curve is a
descriptive tool -- use it alongside analytical judgement and reflexive writing.

---

# Part 5 -- Quality and Output

## Chapter 16: The Members Tab

### What member checking is

Member checking returns a summary of your findings to participants for them to
confirm, correct, or challenge. This strengthens credibility and
trustworthiness by ensuring your interpretations are recognisable to the people
who provided the data.

### Creating a member check

Navigate to the **Members** tab and click **New member check**. Select a
document and a participant label. Optionally restrict to specific codes.

### The check workflow

Each member check tracks a set of coded segments for one participant. Items move
through these statuses:

| Status | Meaning |
|---|---|
| **pending** | No response recorded yet |
| **confirmed** | Participant confirmed this interpretation |
| **disputed** | Participant challenged this interpretation |
| **partial** | Mixed -- some items confirmed, some disputed |

### Recording a response

Click a check row in the table, enter the participant's response text, and
choose a status for each coded item. Click **Save**.

### Exporting a member check

Click **Download** on a check row to export a formatted HTML or plain-text
document listing the coded passages. Send this document to the participant.

---

## Chapter 17: The Journal Tab

### Why the journal matters

Reflexivity -- examining how your own perspective affects the research -- is not
optional in rigorous qualitative work. The **Journal** tab provides a dedicated,
integrated space for this writing. Entries are date-stamped, attributed to the
active coder, categorised by type, searchable, and exportable.

### The five entry types

| Type | When to use it |
|---|---|
| **Analytical** | Interpretive insights, emerging ideas, code development reasoning |
| **Reflexivity** | How your position or experiences may be shaping the analysis |
| **Decision** | A specific analytical or methodological choice and why you made it |
| **Methodological** | Notes on procedures, protocol adaptations, sampling decisions |
| **Other** | Anything that does not fit the above |

### Writing an entry

Navigate to the **Journal** tab. Select an entry type from the sidebar dropdown,
write your entry in the text area, and click **Save entry**.

### Filtering and searching

Use the type filter dropdown and the search box in the card header to find
specific entries -- useful when writing your methods section and retrieving all
decisions about a particular code.

### Exporting

- **Export DOCX/HTML/TXT**: A readable formatted document, suitable for a
  reflexive appendix.
- **Export CSV**: One row per entry with date, type, content, and author.

---

## Chapter 18: The Export Tab

Navigate to the **Export** tab for structured data export:

- **Analytical themes report**: One section per theme with description, linked
  codes, and all associated excerpts. Available as DOCX, HTML, TXT, or JSON.
- **Full codebook export**: All codes with definitions, criteria, memos,
  hierarchy, and colours. Available as DOCX, XLSX, CSV, HTML, or JSON.
- **Raw table export**: Any internal table (codings, codes, sources, cases,
  etc.) as CSV, JSON, or XLSX.

---

## Chapter 19: The Audit Tab

### What the audit log is

The **Audit** tab has two views:

- **Code history**: A complete, append-only record of every change to codes --
  create, update (name, colour, definition, weight, parent), and delete
  operations. Records the timestamp, coder, field changed, old value, and new
  value. This log cannot be edited or deleted.
- **Coding audit**: A log of every coding operation -- when each coded segment
  was created, by whom, and whether it was later deleted.

### Why it matters

The audit log allows you to:

- Report when you introduced a particular code.
- Confirm whether a code definition changed after reliability coding was
  completed.
- Count how many codes you refined between project phases.

Both tables are sortable and filterable by timestamp, coder, and operation type.

---

## Chapter 20: The Help Tab

The **Help** tab contains built-in reference documentation organised by tab and
feature area. Navigate to it at any time to look up how a specific feature works
or what the expected input format is for a given field.

---

# Part 6 -- R API Reference

The R API gives you programmatic access to everything in the project database,
enables automation and batch processing, and connects saturate's data to the
full R ecosystem.

## Chapter 21: Core Project Functions

### `qc_new()` -- Create a project

```{r}
proj <- qc_new(
  path  = "~/studies/my_study.satdb",
  name  = "My Study Title",
  owner = "Researcher Name"
)
```

### `qc_open()` -- Open an existing project

```{r}
proj <- qc_open("~/studies/my_study.satdb")
```

### `qc_close()` -- Close a project

```{r}
qc_close(proj)
```

Always call this when you finish a session.

### `qc_project_info()` -- Read project metadata

```{r}
info <- qc_project_info(proj)
info$name   # project name
info$owner  # owner
```

### `qc_lock_project()` / `qc_unlock_project()` -- Protect a project

Locking prevents any write operations. Use this after finalising a dataset to
guard against accidental changes.

```{r}
qc_lock_project(proj)    # prevent writes
qc_is_locked(proj)       # TRUE
qc_unlock_project(proj)  # re-enable writes
```

### `qc_split_project()` -- Create a contributor copy

Creates a new standalone `.satdb` file containing the codebook, cases, themes,
and a subset (or all) of documents. The contributor file is completely
independent -- it can be coded offline and on a different computer.

```{r}
contrib <- qc_split_project(
  proj,
  path            = "~/collab/contrib_B.satdb",
  source_ids      = c(3L, 7L, 11L),  # NULL copies all documents
  include_codings = FALSE,
  overwrite       = FALSE
)
qc_close(contrib)
```

### `qc_merge_project()` -- Import a contributor's work

```{r}
result <- qc_merge_project(
  proj,
  contributor_path = "~/collab/contrib_B_returned.satdb",
  on_conflict      = "skip",   # or "replace"
  coders           = NULL      # or e.g. c("Researcher B")
)

result$codes_added    # new codes inserted into master
result$sources_added  # new documents inserted into master
result$codings_added  # new codings inserted
result$codings_skip   # duplicate codings skipped
result$themes_added   # new themes inserted
result$memos_added    # new journal entries inserted
```

---

## Chapter 22: Documents

### `qc_import_document()` -- Import a single document

```{r}
qc_import_document(
  proj,
  path        = "~/transcripts/P01_OrgA.txt",
  name        = "P01 -- Organisation A",
  source_type = "interview",
  language    = "en-GB"
)
```

Pass `content = "..."` instead of `path` to import text directly from a
character scalar.

### `qc_import_batch()` -- Batch import from a folder or table

Import an entire directory of text files, or a CSV/TSV/Excel file with one
document per row:

```{r}
# From a directory of .txt files
qc_import_batch(proj, path = "~/transcripts/", format = "dir")

# From a CSV file
qc_import_batch(
  proj,
  path         = "~/data/survey_responses.csv",
  format       = "csv",
  text_col     = "response",
  name_col     = "respondent_id",
  metadata_cols = c("age_group", "region")
)
```

### `qc_list_documents()` -- List all documents

```{r}
docs <- qc_list_documents(proj)
```

Returns columns: `id`, `name`, `source_type`, character count, `n_codings`,
`n_coders`.

### `qc_get_document()` -- Retrieve a single document with its text

```{r}
doc <- qc_get_document(proj, id = 3)
cat(doc$content)
```

### `qc_segment_document()` -- Split a document into segments

```{r}
qc_segment_document(
  proj,
  source_id   = 3,
  method      = "paragraph",
  keep_parent = TRUE
)
```

`method` can be `"paragraph"` (split on blank lines), `"sentence"`, or a custom
regex pattern. The original document is kept by default; set
`keep_parent = FALSE` to replace it with the segments.

### `qc_detect_duplicates()` -- Find near-duplicate documents

```{r}
dups <- qc_detect_duplicates(proj, threshold = 0.85)
```

Returns pairs of documents with Jaccard similarity above the threshold.

### `qc_list_versions()` / `qc_restore_version()` -- Document version history

saturate preserves every version of a document's text when content is edited.

```{r}
versions <- qc_list_versions(proj, source_id = 3)
qc_restore_version(proj, source_id = 3, version_id = 1)
```

---

## Chapter 23: Codes and the Codebook

### `qc_add_code()` -- Create a code

```{r}
new_code <- qc_add_code(
  proj,
  name       = "boundary_blurring",
  color      = "#4A90D9",
  definition = "Subjective experience of work and personal time merging.",
  weight     = 1.0
)
code_id <- new_code$id
```

### `qc_update_code()` -- Update a code

```{r}
qc_update_code(
  proj,
  id         = 7,
  definition = "Updated definition after team discussion on 2026-02-14."
)
```

Only changed fields are written to the audit log.

### `qc_list_codes()` -- List all codes

```{r}
codes <- qc_list_codes(proj)
```

Returns columns including `id`, `name`, `color`, `definition`, `criteria`,
`parent_id`, `depth`, `n_codings`, `categories`, `weight`,
`weight_description`, `level`, `orientation`, `deprecated`.

### `qc_deprecate_code()` / `qc_undeprecate_code()` -- Code lifecycle

```{r}
qc_deprecate_code(proj, code_id = 7, reason = "Merged into isolation_felt.")
qc_undeprecate_code(proj, code_id = 7)
```

Deprecated codes are hidden from the code picker but their codings are
preserved.

### `qc_merge_codes()` -- Merge codes

Reassigns all codings from one or more source codes to a target code, then
soft-deletes the sources:

```{r}
qc_merge_codes(
  proj,
  from_ids = c(12L, 15L),
  into_id  = 7L
)
```

### `qc_split_code()` -- Split a code

Creates a new child code and optionally moves selected codings to it:

```{r}
qc_split_code(
  proj,
  code_id  = 7,
  new_name = "boundary_blurring_temporal",
  color    = "#2ECC71"
)
```

### `qc_add_code_relation()` -- Define a relationship between codes

```{r}
qc_add_code_relation(
  proj,
  code_id_1   = 7,
  code_id_2   = 12,
  relation    = "causes"
)
```

Relations are non-hierarchical (distinct from parent-child) and can carry any
label.

### `qc_snapshot_codebook()` / `qc_diff_snapshots()` -- Versioning

```{r}
qc_snapshot_codebook(proj, name = "Post Phase 1 fieldwork")

diff <- qc_diff_snapshots(
  proj,
  snap_a = "Initial codebook",
  snap_b = "Post Phase 1 fieldwork"
)
```

### `qc_validate_codebook()` -- Validate

```{r}
issues <- qc_validate_codebook(proj)
```

---

## Chapter 24: Codings

### `qc_add_coding()` -- Apply a code to a passage

```{r}
qc_add_coding(
  proj,
  source_id = 3,
  code_id   = 7,
  selfirst  = 1452,
  selast    = 1598,
  coder     = "Researcher A"
)
```

`selfirst` and `selast` are 1-based inclusive character positions.
`substr(content, selfirst, selast)` extracts the exact coded passage.

### `qc_list_codings()` -- List codings

```{r}
all_codings     <- qc_list_codings(proj)
coder_a_codings <- qc_list_codings(proj, coder = "Researcher A")
doc3_codings    <- qc_list_codings(proj, source_id = 3)
```

### `qc_get_coded_segments()` -- Retrieve coded text with context

```{r}
segments <- qc_get_coded_segments(
  proj,
  code_name     = "boundary_blurring",
  context_chars = 100
)
```

### `qc_update_coding_confidence()` -- Update confidence

```{r}
qc_update_coding_confidence(proj, coding_id = 42, confidence = 75)
```

### `qc_uncoded_segments()` -- Find uncoded passages

```{r}
uncoded <- qc_uncoded_segments(proj, source_id = 3)
```

Returns character ranges within the document that have no coding applied.

### `qc_disputed_segments()` -- Find inter-coder disagreements

```{r}
disputed <- qc_disputed_segments(proj, source_id = 3)
```

Returns passages where two or more coders' coding boundaries do not align.

---

## Chapter 25: Annotations and Excerpts

### Annotations

Annotations are position-linked notes attached to a specific character range in
a document, but distinct from codings -- they do not belong to a code and are
not counted in frequency analyses.

```{r}
qc_add_annotation(
  proj,
  source_id = 3,
  selfirst  = 500,
  selast    = 600,
  memo      = "Interesting contradiction with P07's account.",
  coder     = "Researcher A"
)

annotations <- qc_list_annotations(proj, source_id = 3)
```

### Excerpts

Excerpts are labelled passages retained for their illustrative value,
independent of any code.

```{r}
qc_add_excerpt(
  proj,
  source_id = 3,
  selfirst  = 1200,
  selast    = 1350,
  memo      = "Vivid description of the blurred boundary experience."
)

excerpts <- qc_list_excerpts(proj, source_id = 3)
```

---

## Chapter 26: Analysis Functions

### `qc_code_summary()` -- Code frequency summary

```{r}
summary_df <- qc_code_summary(proj)
```

One row per code: total count, documents, coders, mean coding length.

### `qc_code_cooccurrence()` -- Co-occurrence matrix

```{r}
cooc <- qc_code_cooccurrence(proj)
```

Returns a square matrix of code-by-code co-occurrence counts.

### `qc_code_metrics()` -- Dispersion and keyness

```{r}
metrics <- qc_code_metrics(proj)
```

Returns Gries' deviation of proportions (DP) and related metrics per code.

### `qc_code_by_unit()` -- Code frequency by unit

```{r}
by_unit <- qc_code_by_unit(proj, unit = "document")
```

Returns a code-by-document frequency table.

### `qc_cross_tabulate()` -- Codes by case attribute

```{r}
tab <- qc_cross_tabulate(proj, attribute = "working_arrangement")
```

### `qc_temporal_analysis()` -- Coding patterns over time

```{r}
temporal <- qc_temporal_analysis(proj, period = "month")
plot(temporal)
```

### `qc_proximity_query()` -- Find co-occurring passages

```{r}
prox <- qc_proximity_query(
  proj,
  code_id_a = 7,
  code_id_b = 12,
  max_gap   = 200
)
```

Returns codings where code A and code B appear within `max_gap` characters of
each other.

### `qc_triangulate()` -- Tabulate codes against case attributes

```{r}
tri <- qc_triangulate(proj, attribute = "working_arrangement")
```

### `qc_agreement()` -- Pairwise coder agreement

```{r}
agreement <- qc_agreement(
  proj,
  coder_a = "Researcher A",
  coder_b = "Researcher B"
)
```

Returns kappa, percentage agreement, and comparison unit count per code.

### `qc_agreement_matrix()` -- Multi-coder agreement matrix

```{r}
mat <- qc_agreement_matrix(proj)
```

Coder-by-coder matrix of mean kappa values.

### `qc_krippendorff()` -- Krippendorff's alpha

```{r}
alpha <- qc_krippendorff(proj, level = "nominal")
```

Handles multiple coders, missing data, and different measurement levels
(`"nominal"`, `"ordinal"`, `"interval"`). Values above 0.80 are generally
accepted.

### `qc_saturation_curve()` -- Saturation curve

```{r}
curve <- qc_saturation_curve(proj)
plot(curve)
```

---

## Chapter 27: Visualisation Functions

All plot functions return a `ggplot2` object and require `ggplot2` to be
installed.

### `qc_plot_codes()` -- Code frequency bar chart

```{r}
qc_plot_codes(proj, top_n = 20)
```

### `qc_plot_cooccurrence()` -- Co-occurrence heatmap

```{r}
qc_plot_cooccurrence(proj)
```

### `qc_plot_overlap()` -- Coding overlap chart

```{r}
qc_plot_overlap(proj, source_id = 3)
```

Shows how much each pair of codes overlaps within a document.

### `qc_plot_timeline()` -- Coding timeline

```{r}
qc_plot_timeline(proj)
```

Plots coding activity over time.

### `qc_plot_saturation()` -- Saturation curve plot

```{r}
qc_plot_saturation(proj)
```

### `qc_plot_network()` -- Code network graph

```{r}
qc_plot_network(proj)
```

Requires `igraph` and `ggraph`.

### `qc_as_igraph()` -- Export as igraph object

```{r}
g <- qc_as_igraph(proj)
```

Returns the code co-occurrence network as an `igraph` object for custom
network analysis.

### `qc_cb_palette()` -- Colour-blind-safe palette

```{r}
colours <- qc_cb_palette(n = 8, type = "okabe_ito")
```

Returns `n` hex colour codes from a perceptually-distinct, colour-blind-safe
palette. Types: `"okabe_ito"` (default), `"wong"`, `"tol"`.

---

## Chapter 28: Cases, Memos, and Export

### `qc_add_case()` -- Create a case

```{r}
case    <- qc_add_case(proj, name = "P01", memo = "Senior developer, Organisation A.")
case_id <- case$id
```

### `qc_set_case_attribute()` -- Set a case attribute

```{r}
qc_set_case_attribute(
  proj,
  case_id  = 1,
  variable = "working_arrangement",
  value    = "fully remote"
)
```

Setting an attribute that already exists updates the value.

### `qc_case_attributes_wide()` -- Export case attributes

```{r}
attrs <- qc_case_attributes_wide(proj)
```

Returns a wide data frame: one row per case, one column per attribute.

### `qc_add_project_memo()` -- Add a journal entry

```{r}
qc_add_project_memo(
  proj,
  content    = "Ethical approval granted by Faculty Ethics Committee.",
  type       = "methodological",
  created_by = "Researcher A"
)
```

Valid `type` values: `"analytical"`, `"reflexivity"`, `"decision"`,
`"methodological"`, `"other"`.

### `qc_list_project_memos()` -- List journal entries

```{r}
memos <- qc_list_project_memos(proj)
```

### `qc_export_themes_report()` -- Export analytical themes report

```{r}
qc_export_themes_report(
  proj,
  format      = "docx",
  output_path = "~/exports/themes_report.docx"
)
```

Formats: `"docx"`, `"html"`, `"txt"`, `"json"`.

### `qc_export_codebook()` -- Export codebook summary

```{r}
qc_export_codebook(
  proj,
  format               = "xlsx",
  include_definitions  = TRUE,
  include_criteria     = TRUE,
  include_examples     = TRUE,
  n_examples           = 3,
  output_path          = "~/exports/codebook.xlsx"
)
```

Formats: `"docx"`, `"xlsx"`, `"csv"`, `"json"`, `"html"`.

### `qc_export_project_data()` -- Export raw tables

```{r}
qc_export_project_data(
  proj,
  table_name  = "codings",
  format      = "csv",
  output_path = "~/exports/codings.csv"
)
```

### `qc_summary_report()` -- Generate a summary report

```{r}
qc_summary_report(
  proj,
  path   = "~/exports/project_summary.txt",
  format = "text"
)
```

---

# Appendices

## Appendix A: Glossary

**Annotation**: A position-linked note attached to a character range in a
document, distinct from a coding. Annotations are not counted in frequency
analyses.

**Blind mode**: A setting in the Coding tab that hides other coders' highlights,
allowing independent coding.

**Case**: A unit of analysis -- typically a participant, site, or organisation.
Cases can have structured attributes and be linked to documents.

**Category**: A lateral grouping of codes that cuts across the code hierarchy.

**Coder**: Any person who applies codes to passages. Identified by name via
profiles.

**Coding**: The act of attaching a code to a passage of text; also refers to a
specific instance of that attachment.

**Cohen's kappa**: A statistic measuring inter-rater agreement, corrected for
chance. Values: below 0.40 = poor; 0.41-0.60 = moderate; 0.61-0.80 = good;
above 0.80 = very good.

**Contributor file**: A copy of a project created with `qc_split_project()` and
sent to a coder for independent analysis.

**Excerpt**: A labelled passage retained for its illustrative value,
independent of any code.

**Krippendorff's alpha**: A reliability coefficient handling multiple coders,
missing data, and different measurement levels. Values above 0.80 are generally
accepted.

**Master file**: The central project file managed by the lead researcher.

**Member check**: Returning a summary of your findings or interpretation to a
participant for their feedback.

**Merge**: Importing new codes, codings, themes, and memos from a contributor
file into the master project.

**Saturation**: The point at which new data stops producing new codes or
insights.

**selfirst / selast**: 1-based inclusive character positions of a coding.
`substr(content, selfirst, selast)` extracts the coded passage.

**Split**: Creating a standalone copy of a project for a coder to work in
independently.

**Theme**: A higher-level interpretive synthesis of multiple codes and
categories -- an analytical argument or finding.

---

## Appendix B: Keyboard Shortcuts and Workflow Tips

All shortcuts below are active in the **Coding tab** whenever focus is not
inside a text field, dropdown, or open modal. Press **?** at any time to show
the in-app shortcut reference.

### Coding shortcuts

| Key | Action |
|---|---|
| **Enter** | Apply the currently selected code to the selected text |
| **1** -- **9** | Select the Nth code in the picker *and* apply it immediately -- no Enter needed |
| **/** | Focus the code search box |
| **Esc** | Clear the current text selection and dismiss the code picker |

### Navigation shortcuts

| Key | Action |
|---|---|
| **n** | Next uncoded segment within the current document |
| **p** | Previous uncoded segment within the current document |
| **d** | Next disputed or draft segment within the current document |
| **]** | Next document in the list |
| **[** | Previous document in the list |

### Read-aloud shortcuts

| Key | Action |
|---|---|
| **Space** | Play / pause read-aloud (starts from the clicked position, visible scroll position, or beginning) |
| **x** | Stop read-aloud and reset to the beginning |

### View shortcuts

| Key | Action |
|---|---|
| **b** | Toggle blind mode on/off |
| **?** | Open the keyboard shortcuts reference modal |

### Mouse interactions in the document pane

| Interaction | Action |
|---|---|
| Click and drag | Select a text passage; the code picker opens on release |
| Click on a highlight | Open the edit/remove panel for that coding |
| Click anywhere (no drag) | Set the read-aloud start point; a marker line appears at that position |
| Click a line number | Start read-aloud from that line (requires **Show line numbers** to be on in Settings) |
| Select text, then press **Space** | Reads only the selected passage aloud |

### Shortcut scope and limitations

- Shortcuts are suppressed when a text input, dropdown, or SELECT element has
  focus -- type freely in the code search box and memo fields without
  accidentally triggering them.
- Shortcuts are suppressed when any modal dialogue is open.
- The digit shortcuts (**1**--**9**) apply the Nth code from the *filtered*
  list, so typing a few letters in the code search box first narrows the list
  before you press the digit.
- **Space** prevents the default page-scroll behaviour while in the Coding tab;
  use the mouse wheel or trackpad to scroll instead.

### Workflow tips

**Code a document fully before moving to the next.** Reading each document
completely maintains context and reduces the risk of missing important passages.

**Write a journal entry after every analytical session.** Even a two-sentence
note about what felt uncertain or what you decided captures reasoning that is
easy to forget.

**Use provisional codes freely.** You can always merge, rename, or deprecate
codes. A codebook with 60 codes at the end of first-pass coding is normal. The
Validate button identifies unused and orphan codes.

**Run `qc_validate_codebook()` before reliability coding begins.** Missing
definitions and circular hierarchies are sources of coder confusion -- fix them
first.

**Take a codebook snapshot at major milestones.** Before Phase 2, before
submission, and after reviewer feedback are all good moments.

**Back up your `.satdb` file before any bulk auto-code operation.** Auto-code
creates many codings at once; if you apply the wrong pattern, restoring a backup
is faster than removing hundreds of codings.

**Keep attribute names consistent across cases.** `organisation` and `org` are
treated as different variables.

**When using split/merge, keep the master file closed while contributors are
coding.** There is no requirement to do so, but editing the codebook in the
master while a contributor has a copy can cause name conflicts on merge.

**Lock the project after finalising a dataset.** `qc_lock_project()` prevents
accidental writes during analysis or writing-up.

---

## Appendix C: Troubleshooting

### "Cannot open project -- file is already locked"

Another R session has the file open. Find it and run `qc_close(proj)`, or close
that R session. The lock releases automatically when the process terminates;
restart your computer if it persists.

### DuckDB file on a network drive

Network filesystem latency can cause slow performance or intermittent errors.
Work from a local copy and use cloud sync for backup. Never open the file from
two machines simultaneously.

### "Accidentally deleted a code -- can I get it back?"

Yes. saturate never physically deletes data. Restore the soft-deleted code via
the R API:

```{r}
con <- DBI::dbConnect(duckdb::duckdb(), "path/to/study.satdb")
DBI::dbExecute(con, "UPDATE codes SET status = 1 WHERE name = 'your_code_name'")
DBI::dbDisconnect(con)
```

Then reconnect with `qc_open()`. Codings attached to the code can be restored
similarly by updating the `codings` table.

### Import fails -- unsupported file type

Supported extensions are `.txt`, `.docx`, `.pdf`, `.csv`. Convert `.doc`,
`.odt`, or `.rtf` files to `.docx` or `.txt` before importing. For scanned
PDFs, run OCR first.

### The Shiny app shows a blank screen or error

Stop the app (**Ctrl+C**), run `devtools::load_all()` if working from source,
and relaunch with `shiny_saturate(proj)`. Try a different browser if the
problem persists.

### "Maximum upload size exceeded" when merging a contributor file

The default Shiny upload limit is 5 MB. Raise it by passing `max_upload_mb` to
`shiny_saturate()`:

```{r}
shiny_saturate(proj, max_upload_mb = 500L)
```

The default when launching with `shiny_saturate()` is already 500 MB, so this
error should not appear unless your project file is unusually large.

---

*End of saturate User Guide -- Version 0.1.0*
